---
title: "Hands-On_Ex12"
author: "Hei Yee Kai"
date: 11/10/24
date-modified: "last-modified"
execute:
  eval: true
  echo: true
  freeze: true
---

Predictive modelling uses statistical learning or machine learning techniques to predict outcomes. By and large, the event one wants to predict is in the future. However, a set of known outcome and predictors (also known as variables) will be used to calibrate the predictive models.

Geospatial predictive modelling is conceptually rooted in the principle that the occurrences of events being modeled are limited in distribution. When geographically referenced data are used, occurrences of events are neither uniform nor random in distribution over space. There are geospatial factors (infrastructure, sociocultural, topographic, etc.) that constrain and influence where the locations of events occur. Geospatial predictive modeling attempts to describe those constraints and influences by spatially correlating occurrences of historical geospatial locations with environmental factors that represent those constraints and influences.

# Setup

## Packages

We will be using the following packages

```{r}
pacman::p_load(sf, spdep, GWmodel, SpatialML, 
               tmap, rsample, Metrics, tidyverse)
```

## Data

-   Aspatial

    -   HDB Resale data: a list of HDB resale transacted prices in Singapore from Jan 2017 onwards. It is in csv format which can be downloaded from Data.gov.sg.

-   Geospatial

    -   MP14_SUBZONE_WEB_PL: a polygon feature data providing information of URA 2014 Master Plan Planning Subzone boundary data. It is in ESRI shapefile format. This data set was also downloaded from Data.gov.sg

-   Locational factors with geographic coordinates

    -   Downloaded from Data.gov.sg.

        -   Eldercare data is a list of eldercare in Singapore. It is in shapefile format.

        -   Hawker Centre data is a list of hawker centres in Singapore. It is in geojson format.

        -   Parks data is a list of parks in Singapore. It is in geojson format.

        -   Supermarket data is a list of supermarkets in Singapore. It is in geojson format.

        -   CHAS clinics data is a list of CHAS clinics in Singapore. It is in geojson format.

        -   Childcare service data is a list of childcare services in Singapore. It is in geojson format.

        -   Kindergartens data is a list of kindergartens in Singapore. It is in geojson format.

    -   Downloaded from Datamall.lta.gov.sg.

        -   MRT data is a list of MRT/LRT stations in Singapore with the station names and codes. It is in shapefile format.

        -   Bus stops data is a list of bus stops in Singapore. It is in shapefile format.

-   Locational factors without geographic coordinates

    -   Downloaded from Data.gov.sg.
        -   Primary school data is extracted from the list on General information of schools from data.gov portal. It is in csv format.
    -   Retrieved/Scraped from other sources
        -   CBD coordinates obtained from Google.
        -   Shopping malls data is a list of Shopping malls in Singapore obtained from Wikipedia.
        -   Good primary schools is a list of primary schools that are ordered in ranking in terms of popularity and this can be found at Local Salary Forum.

### Import

```{r}
mdata <- read_rds("data/mdata.rds")
```

#### Sampling

The entire data are split into training and test data sets with 65% and 35% respectively by using `initial_split()` of rsample package. rsample is one of the package of tigymodels.

```{r}
#| eval: false
set.seed(1234)
resale_split <- initial_split(mdata, 
                              prop = 6.5/10,)
train_data <- training(resale_split)
test_data <- testing(resale_split)
```

```{r}
#| eval: false
write_rds(train_data, "data/train_data.rds")
write_rds(test_data, "data/test_data.rds")
```

# Correlation Matrix

Before loading the predictors into a predictive model, it is always a good practice to use correlation matrix to examine if there is sign of multicolinearity.

```{r}
mdata_nogeo <- mdata %>%
  st_drop_geometry()
corrplot::corrplot(cor(mdata_nogeo[, 2:17]), 
                   diag = FALSE, 
                   order = "AOE",
                   tl.pos = "td", 
                   tl.cex = 0.5, 
                   method = "number", 
                   type = "upper")
```

The correlation matrix above shows that all the correlation values are below 0.8. Hence, there is no sign of multicolinearity.

# Retrieving the stored data

```{r}
train_data <- read_rds("data/train_data.rds")
test_data <- read_rds("data/test_data.rds")
```

# Building a non-spatial multiple linear regression

```{r}
price_mlr <- lm(resale_price ~ floor_area_sqm +
                  storey_order + remaining_lease_mths +
                  PROX_CBD + PROX_ELDERLYCARE + PROX_HAWKER +
                  PROX_MRT + PROX_PARK + PROX_MALL + 
                  PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                  WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                  WITHIN_1KM_PRISCH,
                data=train_data)
summary(price_mlr)
```

```{r}
#| eval: false
write_rds(price_mlr, "data/price_mlr.rds" ) 
```

# Preparing coordinates data

## Extracting

The code chunk below extract the x,y coordinates of the full, training and test data sets.

```{r}
coords <- st_coordinates(mdata)
coords_train <- st_coordinates(train_data)
coords_test <- st_coordinates(test_data)
```

Before continue, we write all the output into rds for future used.

```{r}
#| eval: false
coords_train <- write_rds(coords_train, "data/coords_train.rds" )
coords_test <- write_rds(coords_test, "data/coords_test.rds" )
```

## Dropping Geometry

First, we will drop geometry column of the sf data.frame by using `st_drop_geometry()` of sf package.

```{r}
train_data <- train_data %>% 
  st_drop_geometry()
```

# Calibrating Random Forest Model

In this section, we will calibrate a model to predict HDB resale price by using random forest function of `ranger` package.

```{r}
#| eval: false
set.seed(1234)
rf <- ranger(resale_price ~ floor_area_sqm + storey_order + 
               remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE + 
               PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL + 
               PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
               WITHIN_350M_CHILDCARE + WITHIN_350M_BUS + 
               WITHIN_1KM_PRISCH,
             data=train_data)
rf
```

```{r}
#| eval: false
write_rds(rf, "data/rf.rds")
```

```{r}
rf <- read_rds("data/rf.rds")
rf
```

# Calibrating Geographical Random Forest Model

In this section, we will calibrate a model to predict HDB resale price by using `grf()` of `SpatialML` package.

## Calibrating using training data

The code chunk below calibrate a geographic random forest model by using `grf()` of `SpatialML` package.

```{r}
#| eval: false
set.seed(1234)
gwRF_adaptive <- grf(formula = resale_price ~ floor_area_sqm + storey_order +
                       remaining_lease_mths + PROX_CBD + PROX_ELDERLYCARE +
                       PROX_HAWKER + PROX_MRT + PROX_PARK + PROX_MALL +
                       PROX_SUPERMARKET + WITHIN_350M_KINDERGARTEN +
                       WITHIN_350M_CHILDCARE + WITHIN_350M_BUS +
                       WITHIN_1KM_PRISCH,
                     dframe=train_data, 
                     bw=55,
                     kernel="adaptive",
                     coords=coords_train)
```

```{r}
#| eval: false
write_rds(gwRF_adaptive, "data/gwRF_adaptive.rds")
```

```{r}
#| eval: false
gwRF_adaptive <- read_rds("data/gwRF_adaptive.rds")
```

## Predicting by using test data

### Preparing the test data

The code chunk below will be used to combine the test data with its corresponding coordinates data.

```{r}
test_data <- cbind(test_data, coords_test) %>%
  st_drop_geometry()
```

### Predicting with test data

Next, `predict.grf()` of `spatialML` package will be used to predict the resale value by using the test data and `gwRF_adaptive` model calibrated earlier.

```{r}
#| eval: false
gwRF_pred <- predict.grf(gwRF_adaptive, 
                           test_data, 
                           x.var.name="X",
                           y.var.name="Y", 
                           local.w=1,
                           global.w=0)
```

Before moving on, let us save the output into rds file for future use.

```{r}
#| eval: false
GRF_pred <- write_rds(gwRF_pred, "data/GRF_pred.rds")
```

### Converting the predicting output into a data frame

The output of the `predict.grf()` is a vector of predicted values. It is wiser to convert it into a data frame for further visualisation and analysis.

```{r}
#| eval: false
GRF_pred <- read_rds("data/GRF_pred.rds")
GRF_pred_df <- as.data.frame(GRF_pred)
```

In the code chunk below, cbind() is used to append the predicted values onto the test_data

```{r}
#| eval: false
test_data_p <- cbind(test_data, GRF_pred_df)
```

```{r}
#| eval: false
write_rds(test_data_p, "data/test_data_p.rds")
```

## Calculating Root Mean Square Error

The root mean square error (RMSE) allows us to measure how far predicted values are from observed values in a regression analysis. In the code chunk below, rmse() of Metrics package is used to compute the RMSE.

```{r}
test_data_p <- read_rds("data/test_data_p.rds")
```

```{r}
rmse(test_data_p$resale_price, 
     test_data_p$GRF_pred)
```

## Visualising the predicted values

Alternatively, scatterplot can be used to visualise the actual resale price and the predicted resale price by using the code chunk below.

```{r}
ggplot(data = test_data_p,
       aes(x = GRF_pred,
           y = resale_price)) +
  geom_point()
```

A better predictive model should have the scatter point close to the diagonal line. The scatter plot can be also used to detect if any outliers in the model.
